---
title: "Peer-graded Assignment: Prediction Assignment Writeup"
author: "Loutfi Kuret"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1) Setting libraries

```{r, results='hide'}
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
set.seed(1337)
```


# 2) Setting and cleaning Input Data 

```{r}
# Download and define the Input datasets 
UrlTraining <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
UrlTesting  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(UrlTraining))
testing  <- read.csv(url(UrlTesting))

inTraining  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainingSet <- training[inTraining, ]
TestingSet  <- training[-inTraining, ]

# remove variables with Nearly Zero Variance
temp <- nearZeroVar(TrainingSet)
TrainingSet <- TrainingSet[, -temp]
TestingSet  <- TestingSet[, -temp]

# remove variables that are mostly NA
temp2    <- sapply(TrainingSet, function(x) mean(is.na(x))) > 0.95
TrainingSet <- TrainingSet[, temp2==FALSE]
TestingSet  <- TestingSet[, temp2==FALSE]

# remove identification variables (columns 1 to 5)
TrainingSet <- TrainingSet[, -(1:5)]
TestingSet  <- TestingSet[, -(1:5)]
```

# 3) Prediction Models

We will compare the accuracy of the Random Forest and Decision Tree.
The model with the best accuracy will be used as prediction on the testing dataset.

# 3.1) Random Forest

```{r}
# Fitting
control <- trainControl(method="cv", number=3, verboseIter=FALSE)
RFfit <- train(classe ~ ., data=TrainingSet, method="rf", trControl=control)
RFfit$finalModel0

# prediction on Test dataset
RF <- predict(RFfit, newdata=TestingSet)
RFmatrix <- confusionMatrix(RF, TestingSet$classe)
RFmatrix

# Confusion matrix and accuracy
plot(RFmatrix$table, col = RFmatrix$byClass, main = paste("Random Forest - Accuracy =", round(RFmatrix$overall['Accuracy'], 4)))
```


# 3.2) Prediction Model Building Method: Decision Trees

```{r}
# fitting
Treefit <- rpart(classe ~ ., data=TrainingSet, method="class")

# prediction on Test dataset
Tree <- predict(Treefit, newdata=TestingSet, type="class")
Treematrix <- confusionMatrix(Tree, TestingSet$classe)
Treematrix

# onfusion matrix and accuracy
plot(Treematrix$table, col = Treematrix$byClass, 
     main = paste("Decision Tree - Accuracy =", round(Treematrix$overall['Accuracy'], 4)))
```


# 4) Test Data Prediction data

The Random Forest model accuracy (0.9973) is higher than the Decision Trees model accuracy (0.7359).

For that reason we will choose the Random Forest model to apply on the testing data.

Below are the results of the predictions

```{r}
final <- predict(RFfit, newdata=testing)
final
```


